---
title: "F21 Natural Selection Data"
output:  prettydoc::html_pretty:
    theme: caymen
    highlight: github
date: "2023-01-31"
editor_options: 
  chunk_output_type: inline
---
## How to recreate some graphs!

For my final project, I decided to recreate some graphs from my BIOL140. They definitely posed a challenge, but I was able to recreate them with some assistance. Thank you Professor Ash!

### Step 1

The original data was not conducive to plotting data in R, so it has to be manipulated.
In an excel file, copy and paste the data points into multiple columns and rename the headers. 
*I decided to rename the the first graph's control data to C1:C4 and the treatment group T1:T2. 
*For the second graph, I named the control group time 1.C:time 6.C and the treatment group time 1.T:time 6.T. 

### Step 2 

Download the data and place it into your BIOL1007 folder labeled "Data". (It could be saved elsewhere if you would like)

### Step 3

Load the data into R using "read.table" and getting graphing! The steps I used are used below.

```{r}
### Graph 1

#download the necessary packages
library(tidyverse)
library(dplyr)
library(ggthemes)
library(tidyr)

NSData <- read.table(file="Data/New_F21_Natural_Selection.csv", header=TRUE, sep= ",") #downloaded data and put into my data folder in BIOL1007
#View(NSData)
glimpse(NSData)

NSData$T4 <- as.numeric(NSData$T4) #This line was showing up as a character, so set it to numeric
glimpse(NSData)


NSData1 = na.omit(NSData) #removed NAs
glimpse(NSData1)
#View(NSData1)

NSD <- NSData1 %>% #reorganized data layout
  pivot_longer(
    cols = 1:8, 
    names_to = "Group", 
    values_to = "Zone_of_Inhibiton", 
    values_drop_na = TRUE)
#View(NSD)

library(plotrix) #helpful when plotting because it has various labeling, axis, and color scaling functions

C<- NSD %>%
  group_by(Group)%>%
  summarize(average_zoi=mean(Zone_of_Inhibiton),se_zoi= std.error(Zone_of_Inhibiton)) %>%
  mutate(treatment= rep(c("Control", "Treatment"), each= 4), Round= rep(1:4, 2)) #used piping to end up if data organized by round and treatment group

#View(C)
df<- data.frame(C) #maade a new data frame
df

#plotting data in scatterplot
 sp<- ggplot(data=df, aes(x= Round, y=average_zoi, color=treatment)) + #color=treatment plots data from the treatment column we just made
  geom_point() +
  geom_line(size=1) +
   geom_errorbar(aes(ymin=average_zoi-se_zoi, ymax=average_zoi+se_zoi), width=0.1) + #graohs the error bars
  ylab("Zone of Inhibiton (mm)") + 
  xlab("Rounds") +
  theme(legend.position = "bottom") +
  theme(legend.key.size = unit(0.2, 'cm')) +
  theme(legend.title=element_blank()) +
  theme(axis.line = element_line(color='grey'), #manual way to get the x/y axis lines I wanted
    plot.background = element_blank(),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_rect(fill = 'white', color = 'white')) +
   scale_y_continuous(breaks= c(0, 1, 2, 3, 4, 5, 6, 7, 8), limits=c(0, 8)) + #manual way to get the axis tics of your choosing
  theme(axis.title.y = element_text(size=8,
      colour = "black")) +
  theme(axis.title.x = element_text(size=8,
      colour = "black"))
  sp1<- sp
  sp1 + scale_color_manual(values=c("#CC6600","#0666CC")) #changes the colors


```

```{r}
### Graph 2

A600Data <- read.table(file="Data/A600_data_Updated3.csv", header=TRUE, sep= ",") #downloaded data and put into my data folder in BIOL1007
#View(A600Data)

A600Data$time.4.T <- as.numeric(A600Data$time.4.T) #This line was showing up as a character, so set it to numeric
glimpse(A600Data)

A600Data %>%
   mutate_all(as.character) #getting error saying time.1.C <double> and time.4.T is <character> // doesn't work lol
is.numeric(A600Data$time.2.T) # same as above

A600 <- A600Data %>%
  pivot_longer(
    cols = starts_with("time"), #specify which columns you want to make 'longer'
    names_to = "Time", #name of new column of header names
    values_to = "A600", #name of new column of cell values
    values_drop_na = TRUE
  )
##changed the way the data was set up
#View(A600)

A600$Trt <- ifelse(grepl("C", A600$Time), "Control", "Treatment")
#used the ifelse with grepl to organize the data // grepl identifies specific substrings(or characteristics) and finds all data with the same pattern


#case_when using grep ~
A6<- A600 %>%
  mutate(Group = case_when(grepl("1", Time) ~ '1',
                          grepl("2", Time)  ~ '2',
                          grepl("3", Time)  ~ '3',
                          grepl("4", Time)  ~ '4',
                          grepl("5", Time)  ~ '5',
                          grepl("6", Time)  ~ '6'))
  ##case_when tests for a specific condition and then gives a value if the condition is TRUE (kinda like if)
##used in conjunction with grepl to organize the data by time increments 

d<- A6 %>%
  group_by(Time, Trt)%>%
  summarize(average_absorb=mean(A600),se_absorb= std.error(A600))
  d$minutes <- factor(rep(c("0", "30", "60", "90", "120", "150"), each= 2), levels= c("0", "30", "60", "90", "120", "150")) #needed to add new minutes column
#piped data to find mean and SE 

#View(d)

#plotting data in bar grpah 

bp<- ggplot(data=d, aes(x=factor(minutes), y= average_absorb, fill= Trt)) + #fill is used to put the data you want into the graph
   geom_col(position="dodge") + #organizes data 
  geom_errorbar(aes(ymin=average_absorb-se_absorb, ymax=average_absorb+se_absorb), width= 0.1, position = position_dodge(width=0.85)) + #added error bars
  ylab("A600 (Absorbance at 600nm)") + 
  xlab("Time (minutes)") +
  scale_x_discrete(labels=c("0", "30", "60", "90","120","150")) + #removed time.1.C... from the bottom
  scale_y_continuous(breaks= c(0, .1,.2,.3,.4,.5,.6), limits=c(0, .6)) + #manual axis tic increments 
  scale_fill_manual(values=c("#0666CC","#CC6600")) + #chaging the colors
  theme(legend.position = "bottom") +
  theme(legend.key.size = unit(0.2, 'cm')) + #changed legend size
  theme(legend.title=element_blank()) + #removed legend title
  theme(axis.line = element_line(color='grey'), #manual way to get the x/y axis lines I wanted
    plot.background = element_blank(),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_rect(fill = 'white', color = 'white')) +
  theme(axis.title.y = element_text(size=8,
      colour = "black")) +
  theme(axis.title.x = element_text(size=8,
      colour = "black"))

  bp
  
  

 

```

### Great Job!